---
title: STA 529 2.0 Data Mining
author: Dr Thiyanga S. Talagala
date: "AdaBoost and Gradient Boosting"
institute: "Lecture 4"
output: binb::metropolis
fontsize: 12pt
header-includes: 
  - \usepackage{multirow}
---

```{r,setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```


## Random Forests vs AdaBoost


\begin{table}[!h]
\begin{tabular}{|p{2cm}|p{3cm}|p{4cm}|}
\hline
Category & Random Forest  & AdaBoost \\ \hline
Tree size & Fully grown & Decision stump \\ \hline
Tree construction & Independent & Each stump is made by taking the previous stump’s mistakes into account \\ \hline
Say for each tree & Equal & Some stumps get more say \\ \hline
 & parallel  &  sequential\\ \hline
\end{tabular}
\end{table}

## Data

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=NA}
smoking <- c("Yes", "No", "Yes", "Yes", "No", "No", "Yes", "Yes")
family_history <- c("Yes", "Yes", "No", "Yes", "Yes", "Yes", "No", "Yes")
height <- c(5.3, 6, 5.2, 5, 5.9, 4.7, 5.2, 5.4)
cancer <- c("Yes", "Yes", "Yes", "Yes", "No", "No", "No", "No")
df <- data.frame(smoking, family_history, height, cancer)
df
```

## Creating the first decision stump


1) Initial weights

```{r, comment=NA, echo=FALSE, size="tiny"}
df$Ini_weights <- rep(0.125, 8)
df
```

## Decision stump  (In-class)

Smoking ?

Height? Splitting point?

Family_history?


## Decision stump

Your turn

Compute node Gini coefficient for Smoking and Family History and Height > 5.1

## Frist decision stump

In-class


Compute total error.

## Amount of say

$$\text{Amount of say} = \frac{1}{2} log \left(\frac{1-\text{Total Error}}{\text{Total Error}}\right)$$

Compute the amount of say for the first decision stump.

## Amount of say vs Total error

```{r, echo=FALSE}
library(ggplot2)
x <- seq(0.001, 0.999, length = 100)
y <- 0.5*(log((1-x)/x))
df2 <- data.frame(total_error=x, amount_of_say=y)
ggplot(df2, aes(x=total_error, y=amount_of_say)) + geom_line()
```

## Building the next decision stump

```{r, echo=FALSE, comment=NA}
df
```

## New observation weights for incorrect classifications

$$\text{New observation weight} = \text{initial weight} \times e^{\text{Amount of say}}$$

```{r, echo=FALSE}
x <- seq(1, 10, length = 100)
y <- exp(x)
df2 <- data.frame(amount_of_say=x, e_amount_of_say=y)
ggplot(df2, aes(x=amount_of_say, y=e_amount_of_say)) + geom_line()
```


## New observation weights for correct classifications

$$\text{New observation weight} = \text{initial weight} \times e^{\text{-Amount of say}}$$

```{r, echo=FALSE}
x <- seq(1, 10, length = 100)
y <- exp(-x)
df2 <- data.frame(amount_of_say=x, e_amount_of_say=y)
ggplot(df2, aes(x=amount_of_say, y=e_amount_of_say)) + geom_line()
```

## Initial weights + new weights

```{r, echo=FALSE, comment=NA}
int_weights <- rep(0.125, 8)
new_weights <- c(0.05, 0.05, 0.05, 0.33, 0.05, 0.05, 0.05, 0.05)
df4 <- data.frame(ini_weights=int_weights, new_weights = new_weights)
df4
```

## Initial weights + new weights + normalized weights

```{r, echo=FALSE, comment=NA}
int_weights <- rep(0.125, 8)
new_weights <- c(0.05, 0.05, 0.05, 0.33, 0.05, 0.05, 0.05, 0.05)
normalized_weights <- new_weights/0.68
df4 <- data.frame(ini_weights=int_weights, new_weights = new_weights, normalized_weights=normalized_weights)
df4
```

## Create the next decision stump - inclass



## Adaboost final decision - inclass


## Adaboost vs Gradient Boosting Algorithm



\begin{table}[!h]
\begin{tabular}{|p{2cm}|p{3cm}|p{4cm}|}
\hline
Category & AdaBoost  & Gradient Boosting \\ \hline
Tree size & Decision stump & Statrts by making a single leaf and then grow trees\\ \hline
Tree construction & Up weights the observations misclassified before & identify observations by large residuals computed in the previous iteration  \\ \hline
Say for each tree & Not Equal & Equal \\ \hline
Tree construction & sequential  &  sequential\\ \hline
\end{tabular}
\end{table}

# Gradient Boosting - Regression

## Dataset

```{r, echo=FALSE, comment=NA}
colour <- c("blue", "brown", "blue", "black", "brown", "blue")
gender <- c("M", "F", "F", "M", "M", "F")
tusk_length <- c(1.6, 1.6, 1.5, 1.8, 1.5, 1.4)
bmi <- c(2.3, 2.4, 3.2, 3.4, 3.5, 2.1)
df4 <- data.frame(colour = colour, gender = gender, tusk_length = tusk_length, bmi)
df4
```

## Initial guess - inclass


## Compute residuals - inclass


## Predict BMI - inclass

## Continue building trees - inclass

## Final prediction


