---
title: " Support Vector Machines (SVM)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

The objective of the support vector machine algorithm is to find a hyperplane in an N-dimensional space(N — the number of features) that distinctly classifies the data points.

# Hyperplanes and Support Vectors

In-class

# Linearly separable vs non-linearly separable

In-class

\vspace{5cm}

# Illustration

- Maximized margins

```{r, echo=FALSE, message=FALSE,warning=FALSE}
library(tidyverse)
set.seed(12)
x <- c(rnorm(20, mean = 2), rnorm(20, mean=5))
y1 <- 0.5*x[1:20] + rnorm(20)
y2 <- 0.5*x[21:40] + rnorm(20)
y <- c(y1, y2)
z <- factor(c(rep(1, 20), rep(0, 20)))
df <- tibble(x=x, y=y, z=z)
ggplot(df, aes(x=x, y=y,col=z)) + geom_point(size=2)
```


- There are many possible hyperplanes that could be chosen to classify data points.

- Our objective is to find a plane that has the maximum margin.

- Support vectors: data points that are closer to the hyperplane and influence the position and orientation of the hyperplane.

# R code

```{r}
library(palmerpenguins)
data(penguins)

```

Splitting data into training and test set

```{r}
library(tidymodels)
set.seed(123)
penguins <- na.omit(penguins) 
penguins_split <- initial_split(penguins, strata = species)
penguins_train <- training(penguins_split)
summary(penguins_train)
penguins_test <- testing(penguins_split)
summary(penguins_test)

```


Next, let’s create cross-validation resamples of the training data, to evaluate our models.

```{r}
set.seed(234)
penguins_folds <- vfold_cv(penguins_train, strata = species)
penguins_folds
```

Next, let’s preprocess our data to get it ready for modeling.

```{r}
penguins_rec <- recipe(species ~ island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g + sex, data = penguins_train) 

penguins_rec2 <- recipe(species ~ island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g + sex, data = penguins_train)  %>%
   step_normalize(body_mass_g)

penguins_prep <- prep(penguins_rec)
penguins_prep
juice(penguins_prep)
```


## Random forest

```{r}
rf_spec <- rand_forest(trees = 1000) %>%
  set_engine("ranger") %>%
  set_mode("classification")

rf_spec
```

## SVM

```{r}
svm_spec <- svm_rbf(cost = 0.5) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

svm_spec
```


```{r}
penguins_wf <- workflow() %>%
  add_recipe(penguins_rec)

penguins_wf
```

```{r}
## Illustration only

penguins_prep2 <- prep(penguins_rec2)
penguins_prep2
juice(penguins_prep2)
penguins_wf2 <- workflow() %>%
  add_recipe(penguins_rec2)
penguins_wf2
```

Now we can add a model, and the fit to each of the resamples. First, we can fit the random forest model.

```{r}
set.seed(1234)
rf_rs <- penguins_wf %>%
  add_model(rf_spec) %>%
  fit_resamples(
    resamples = penguins_folds,
    metrics = metric_set(roc_auc, accuracy, sens, spec),
    control = control_grid(save_pred = TRUE)
  )
rf_rs 

```

## SVM

```{r}
svm_rs <- penguins_wf %>%
  add_model(svm_spec) %>%
  fit_resamples(
    resamples = penguins_folds,
    metrics = metric_set(roc_auc, accuracy, sens, spec),
    control = control_grid(save_pred = TRUE)
  )

```

## Evaluate Model

Random forests

```{r}
collect_metrics(rf_rs)
conf_mat_resampled(rf_rs)
```

SVM

```{r}
collect_metrics(svm_rs)
conf_mat_resampled(svm_rs)
```

## Evaluation on test set

```{r}
rf_fit <- rf_spec %>%
  fit(species ~ island + bill_length_mm + bill_depth_mm + flipper_length_mm + body_mass_g + sex, data = penguins_train)
rf_fit
rf_testing_pred <- 
  predict(rf_fit, penguins_test, type = "prob")
rf_testing_pred
```

```{r}
rf_training_pred <- 
  predict(rf_fit, penguins_train) %>% 
  bind_cols(predict(rf_fit, penguins_train, type = "prob")) %>% 
  # Add the true outcome data back in
  bind_cols(penguins_train %>% 
              select(species))
library(yardstick)
rf_training_pred %>%                # training set predictions
  accuracy(truth = species, .pred_class)
```




"Resampling allows us to simulate how well our model will perform on new data, and the test set acts as the final, unbiased check for our model’s performance."

Source: https://www.tidymodels.org/start/resampling/

